---
title: Coronavirus in Saudi Arabia
author: Emily Sheehan
date: '2020-09-03'
slug: coronavirus-in-saudi-arabia
categories: []
tags: []
subtitle: 'Comparison of the incidence of COVID-19 and MERS-COV in Saudi Arabia'
summary: 'Comparison of the incidence of COVID-19 and MERS-COV in Saudi Arabia'
authors: []
lastmod: '2020-09-03T14:35:54+10:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r load-libraries, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(coronavirus)
library(widgetframe)
library(kableExtra)
```

```{r load-data}
data("coronavirus")
```

```{r filter-saudi}
saudi <- coronavirus %>%
  filter(country == "Saudi Arabia") %>%
  select(c("date",
           "type",
           "cases")) 
```

## Data Description

The coronavirus data comprises of 7 variables and over 175,000 observations for a number of countries. The variables include; the date, province, country, latitude, longitude, type of case (i.e. confirmed/recovered) and number of daily cases. 

The raw data has been collected on a daily basis since the 22nd of January 2020 by John Hopkins University Center for Systems Science and Engineering. The sources for the raw data include; the World Health Organisation, the National Health Commission of the People’s Republic of China, the European Center for Disease Prevention and Control, the U.S Centers for Disease Control and Prevention, local media reports, 1point3acres, Worldometers.info, BNO, DXY and other state and national government health departments.

This raw data has been tidied and collated into a package titled _coronavirus_, which has been used for this analysis. 

## An Explanation 

Coronavirus is the name given to a range of respiratory illnesses originating from the same family (Raoult et al., 2020). Each virus varys in its symptoms, with some strains being more fatal than others.Some of the viruses include; Middle Eastern Respiratory Syndrome (MERS-CoV), Severe acute respiratory syndrome (SARS-CoV) and COVID-19. 

The Middle Eastern Respiratory Syndrome was first identified in Saudi Arabia in 2012. This analysis hopes to provide insight and compare the infection rate of MERS-CoV and COVID-19. 

## A Comparison

Since appearing in Saudi Arabia in 2012, there have been a total of 2494 confirmed cases of MERS-CoV worldwide. Saudi Arabia was home to around 2102 infections, and had 780 deaths.

Comparatively, there have been well **over** 317,000 cases of COVID-19 in Saudi Arabia since early 2020. The summary tables below provide some insight into the new cases and recovered cases. The maximum number of new cases on any given day was 867 cases, which is over 40% of the total number of MERS-CoV infections in Saudi Arabia over 8 years. The average number of new infections is around 4.283784 cases per day, while the average number of recovered cases per day is 5.085586. The maximum number of recovered cases on a given day was 5014 cases. 

Ultimately, the summary statistics and the interactive table suggest a clear difference in the incidence of MERS-CoV and COVID-19. Although MERS-CoV is much more fatal with an alarmingly high death rate, COVID-19 is much more infectious. 


```{r saudi-confirmed}
saudi_confirmed <- coronavirus %>%
  filter(country == "Saudi Arabia",
         type == "confirmed") %>%
  mutate(previous_day_confirmed = lag(cases),
         new_confirmed = cases-previous_day_confirmed,
         total_confirmed = sum(cases)) %>%
  filter(date != "2020-01-22") %>%
  select(-country,
         -lat,
         -long,
         -type,
         -province) %>%
  summarise(average_new_cases = mean(new_confirmed),
            max_new_cases = max(new_confirmed),
            total = sum(cases)) %>%
  kable(col.names = c("Average New Cases per day", "Maximum Number of New Cases per day", "Total Number of Cases"), 
        caption = "Summary Statistics for New COVID-19 Cases in Saudi Arabia") %>%
  kable_styling()
saudi_confirmed
```

```{r saudi-recovered}
saudi_recovered <- coronavirus %>%
  filter(country == "Saudi Arabia",
         type == "recovered") %>%
  mutate(previous_day_recovered = lag(cases),
         new_recovered_cases = cases-previous_day_recovered,
         total_recovered = sum(cases)) %>%
  filter(date != "2020-01-22") %>%
  select(-country,
         -lat,
         -long,
         -type,
         -province) %>%
  summarise(average_recovered_cases = mean(new_recovered_cases),
            max_recovered_cases = max(new_recovered_cases),
            total = sum(cases)) %>%
  kable(col.names = c("Average Recovered Cases per day", "Maximum Number of Recovered Cases per day", "Total Number of Recovered Cases"), 
        caption = "Summary Statistics for Recovered COVID-19 Cases in Saudi Arabia") %>%
  kable_styling()
saudi_recovered
```

```{r interactive-table-saudi}
interactive_saudi <- DT::datatable(saudi,
              colnames = c("Date",
                           "Confirmed/Recovered",
                           "Number of Cases"),
              caption = "Number of Confirmed or Recovered Coronavirus Cases on a given Date in Saudi Arabia", 
              options = list(pageLength = 8))
widgetframe::frameWidget(interactive_saudi)
```


## References 

Bhaskar Karambelkar (2017). widgetframe: 'Htmlwidgets' in Responsive 'iframes'. R package version 0.3.1.
  https://CRAN.R-project.org/package=widgetframe
  
Hao Zhu (2020). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax. R package version 1.2.1.
  https://CRAN.R-project.org/package=kableExtra

Middle East respiratory syndrome coronavirus (MERS-CoV). (2020, March 10). Retrieved 4 September 2020, from   
  https://www.who.int/emergencies/mers-cov/en/  

Rami Krispin and Jarrett Byrnes (2020). coronavirus: The 2019 Novel Coronavirus COVID-19 (2019-nCoV) Dataset.
  R package version 0.3.0.9000. https://github.com/RamiKrispin/coronavirus
  
Raoult, D., Zumla, A., Locatelli, F., Ippolito, G., & Kroemer, G. (2020). Coronavirus infections: Epidemiological, clinical
  and immunological features and hypotheses. Cell Stress & Chaperones, 4(4), 66–75.
  
Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686,
  https://doi.org/10.21105/joss.01686

