---
title: Coronavirus in Saudi Arabia
author: Emily Sheehan
date: '2020-09-02'
slug: coronavirus-in-saudi-arabia
categories: []
tags: []
subtitle: 'Comparison of the incidence of COVID-19 and MERS-COV in Saudi Arabia'
summary: 'Comparison of the incidence of COVID-19 and MERS-COV in Saudi Arabia'
authors: []
lastmod: '2020-09-02T10:37:51+10:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r load-libraries, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(coronavirus)
library(widgetframe)
```

```{r load-data}
data("coronavirus")
```

```{r filter-saudi}
saudi <- coronavirus %>%
  filter(country == "Saudi Arabia") %>%
  select(c("date",
           "type",
           "cases")) 
```

# Data Description

The coronavirus data comprises of 7 variables and over 175,000 observations for a number of countries. The variables include; the date, province, country, latitude, longitude, type of case (i.e. confirmed/recovered) and number of daily cases. 

The raw data has been collected by the John Hopkins University Centre for Systems Science and Engineering on a daily basis since the 22nd of January 2020. This data has been tidied and collated into a package titled _coronavirus_, which has been used for this analysis. 


```{r interactive-table-saudi}
interactive_saudi <- DT::datatable(saudi,
              colnames = c("Date",
                           "Confirmed/Recovered",
                           "Number of Cases"),
              caption = "Number of Confirmed or Recovered Coronavirus Cases on a given Date in Saudi Arabia", 
              options = list(pageLength = 4))
widgetframe::frameWidget(interactive_saudi)
```



```{r}
saudi_confirmed <- coronavirus %>%
  filter(country == "Saudi Arabia",
         type == "confirmed") %>%
  mutate(previous_day_confirmed = lag(cases),
         new_confirmed = cases-previous_day_confirmed,
         total_confirmed = sum(cases)) %>%
  select(-country,
         -lat,
         -long,
         -type,
         -province)
```

```{r}
saudi_recovered <- coronavirus %>%
  filter(country == "Saudi Arabia",
         type == "recovered") %>%
  mutate(previous_day_recovered = lag(cases),
         new_recovered_cases = cases-previous_day_recovered,
         total_recovered = sum(cases)) %>%
  select(-country,
         -lat,
         -long,
         -type,
         -province)
```


